---
title: "Music Behavior Around The World"
author: "Benjamin Rim, Pratham Kataria, Christian Kjaer, Paige Stanley"
date: "11/14/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE,  warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Important values

``` {r  echo = FALSE, results = 'asis',  warning=FALSE}
spotify_dataset <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-paigestanley1/main/data/spotify_dataset.csv")
artists <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/top_artists.csv")
songs_normalize <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-Chkjaer/main/data/songs_normalize.csv")
 
summary_info <- list()
summary_info$artist_country_freq <- names(which.max(table(artists$country)))
summary_info$song_year_freq <- names(which.max(table(songs_normalize$year)))
summary_info$genre_freq <- names(which.max(table(songs_normalize$genre)))
summary_info$num_countries <- length(unique(artists$country))
summary_info$max_followers <- names(which.max(table(spotify_dataset$Artist.Followers)))

```

After creating these charts, we can see the beginnings of answers to our research questions. Through examing our data sets closer we have found trends in where the most top artists originate from, the top genres, as well as trends of music through time. The majority of top artists are from the `r summary_info$artists_country_freq`. However, there are `r summary_info$num_countries` countries represented in our data set. The most popular genre within our data sets is `r summary_info$genre_freq` and the year with the most songs that became top songs was `r summary_info$song_year_freq`. We are starting to see the trend of how the advancement of Spotify and the internet in recent years has impacted listening patterns. One way we are tracking the popularity of artists is through their Spotify follower counts. The artist with the most followers is Taylor Swift with `r summary_info$max_followers` followers. All of these values are from the most recent year we have in our data sets, 2021. 

## Table

```{r  echo = FALSE, results = 'asis',  warning=FALSE, message=FALSE}
library("dplyr")
library("ggplot2")
library("plotly")

spotify_dataset <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-paigestanley1/main/data/spotify_dataset.csv")
artists <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/top_artists.csv")
songs_normalize <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-Chkjaer/main/data/songs_normalize.csv")

artists_table <- artists %>%
  group_by(artist_mb) %>%
  select(artist_mb, country, tags_mb, listeners_lastfm)

artists_table <- artists_table %>%
  rename(artist = artist_mb) %>%
  rename(genre = tags_mb) %>%
  rename(country_of_origin = country) %>%
  rename(listeners = listeners_lastfm)

table <- left_join(artists_table, songs_normalize, by = "artist")

spotify_dataset <- spotify_dataset %>%
  rename(artist = Artist)

table_final <- left_join(table, spotify_dataset, by = "artist")

table_final <- table_final %>%
  select(artist, country_of_origin, listeners, popularity, song, year, genre.x, Artist.Followers)

table_final <- table_final[!duplicated(table_final$song), ]

table_final <- table_final %>%
  rename(genre = genre.x) %>%
  rename(artist_followers = Artist.Followers)

table_final <- table_final %>%
  select(artist, country_of_origin, listeners, song, year, genre, artist_followers)

table_final <- table_final[!duplicated(table_final$artist), ]

library(knitr)
kable(table_final[1:25, ], caption = "Aggregate Data Table")

```

### Summary Paragraph

This table shows aggregate data from all three of our charts. The table is grouped by artist, where it shows the top Spotify artists, where they are from, their genre, the total amount of listeners they have on Spotify, the number of followers they have on Spotify, and one of their top songs from 2000-2019, and when each song came out. Through these columns, the table displays which genres are the most popular, which artists are the most popular and where they are from, as well as which songs are they most popular. It is clear through the table that the most popular artists are from the US and UK and that the most popular genres are pop and rock. This table also answers the question that while Spotify is used worldwide, the majority of its popular music is from the US and UK. We wonder if this is because of Spotify's recommendation algorithm or if English music is truly the most widely enjoyed

## Chart 1

```{r  echo = FALSE, results = 'asis',  warning=FALSE, message=FALSE}
library("tidyverse")
library("stringr")
library("ggplot2")

top_spotify_hits2000_2019 <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-Chkjaer/main/data/songs_normalize.csv", header = TRUE, stringsAsFactors = FALSE)

all_genres <- as.list(unique(unlist(strsplit(top_spotify_hits2000_2019$genre, ", "))))

create_df <- function(the_genre) {
  new_data <- top_spotify_hits2000_2019 %>%
    filter(str_detect(genre, the_genre)) %>%
    group_by(year) %>%
    summarise(total_popularity_of_genre = sum(popularity, na.rm = TRUE)) %>%
    return(new_data)
}

create_plot_for_genre <- function(new_data) {
  ggplot(new_data, aes(x = year, y = total_popularity_of_genre)) + 
    geom_point() +
    labs(title = "Popularity of Rock from 2000 - 2019",
         x = "Year",
         y = "Total Popularity of Rock")
}

rock_df <- create_df("rock")
create_plot_for_genre(rock_df)

for (genres in all_genres) {
  df <- create_df(genres)
  create_plot_for_genre(df)
}

```

This is an example of how the popularity of rock has decreased over time. You can use the code above to filter select genres such as rock, pop, R&B, etc. to see the popularity over time. Each dot on the chart represents how popular the genre was on the scale used in our data set.

## Chart 2

```{r  echo = FALSE, results = 'asis',  warning=FALSE, message=FALSE}
library("tidyverse")
library("ggplot2")

  artists <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/top_artists.csv", stringsAsFactors = FALSE)
  capitals <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/concap.csv")
top_artists <- artists %>% 
  arrange(-listeners_lastfm) %>%
  head(1000) %>% rename(country = country)

capitals <- rename(capitals, country = CountryName)
  
top_artists <- top_artists %>%
  left_join(capitals, by = "country")

country_shape <- map_data("world")

view(country_shape)

world_map <- ggplot(country_shape) +
  geom_polygon(
    mapping = aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "black",
    size = .01,
  ) +
  coord_map() +
  labs(
    title = "Top 1000 Spotify Artists' Home Countries by Capital and Amount of Listeners",
    x = NULL,
    y = NULL,
  )

world_map + 
  geom_point(
    data = top_artists,
    mapping = aes(x = CapitalLongitude, y = CapitalLatitude, size = listeners_lastfm),
    color = "green",
    alpha = .3
  ) + 
  labs(
    size = "Listeners"
  )
  
```

This chart shows a map of the world where each green dot represents an artist, and is placed on the capital of the artists home country. The size of the dot represents how many Spotify listeners that artist has.