---
title: "Music Behavior Around The World"
author: "Benjamin Rim, Pratham Kataria, Christian Kjaer, Paige Stanley"
date: "11/14/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE,  warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dynamic Paragraph

After creating these three charts, we can see the beginnings of answers to our research questions. In Chart 1, the values we see are community and upliftment. There are clear winners of which genres are the most popular so we see that people enjoy to listen to the same music as the people around them. The clear winners are also upbeat genres like pop, rock, and rap. In Chart 2 we see the value of trends and social conformity. None of the genres had a straight line in them which shows that there are trends of which genres are popular each year. That means that people follow others' interests in music. Finally, the third chart shows the value of culture. The world map shows the amount of Spotify listeners in each country. As we can see through the table, most of Spotify's top artists are from the US, so through the chart we can see where American and UK music is the most popular vs where countries may be listening to other kinds of music that may be offered through other streaming services.

## Table

```{r echo = FALSE, results = 'asis',  warning=FALSE, message = FALSE}
library("dplyr")
library("ggplot2")
library("plotly")

spotify_dataset <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-paigestanley1/main/data/spotify_dataset.csv")
artists <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/top_artists.csv")
songs_normalize <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-Chkjaer/main/data/songs_normalize.csv")

artists_table <- artists %>%
  group_by(artist_mb) %>%
  select(artist_mb, country, tags_mb, listeners_lastfm)

artists_table <- artists_table %>%
  rename(artist = artist_mb) %>%
  rename(genre = tags_mb) %>%
  rename(country_of_origin = country) %>%
  rename(listeners = listeners_lastfm)

table <- left_join(artists_table, songs_normalize, by = "artist")

spotify_dataset <- spotify_dataset %>%
  rename(artist = Artist)

table_final <- left_join(table, spotify_dataset, by = "artist")

table_final <- table_final %>%
  select(artist, country_of_origin, listeners, popularity, song, year, genre.x, Artist.Followers)

table_final <- table_final[!duplicated(table_final$song), ]

table_final <- table_final %>%
  rename(genre = genre.x) %>%
  rename(artist_followers = Artist.Followers)

table_final <- table_final %>%
  select(artist, country_of_origin, listeners, song, year, genre, artist_followers)

table_final <- table_final[!duplicated(table_final$artist), ]

library(knitr)
kable(table_final[1:25, ], caption = "Aggregate Data Table")

```

### Summary Paragraph

This table shows aggregate data from all three of our charts. The table is grouped by artist, where it shows the top Spotify artists, where they are from, their genre, the total amount of listeners they have on Spotify, the number of followers they have on Spotify, and one of their top songs from 2000-2019, and when each song came out. Through these columns, the table displays which genres are the most popular, which artists are the most popular and where they are from, as well as which songs are they most popular. It is clear through the table that the most popular artists are from the US and UK and that the most popular genres are pop and rock. This table also answers the question that while Spotify is used worldwide, the majority of its popular music is from the US and UK. We wonder if this is because of Spotify's recommendation algorithm or if English music is truly the most widely enjoyed

## Chart 1

```{r echo = FALSE, results = 'asis',  warning=FALSE, message = FALSE}
library("ggplot2")
library("dplyr")

spotify <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-paigestanley1/main/data/spotify_dataset.csv")

stream <- spotify%>%
  arrange(desc(Streams))%>%
  head(10)

stream %>%
  ggplot(aes(x = Streams,
            y = reorder(Genre, Streams))) +
 geom_col(width= .3, color = "black") +
  labs(title = "Most Streamed Genres",
      x = "Streams",
      y = "Genre")

```

This chart shows the most popular genres on spotify. The top is pop while the bottom is jazz. It also shows the number of streams for each genre on the bottom.

## Chart 2

```{r echo = FALSE, results = 'asis',  warning=FALSE, message=FALSE}
library("tidyverse")
library("stringr")
library("ggplot2")

top_spotify_hits2000_2019 <- read.csv("https://raw.githubusercontent.com/info201b-au2022/project-Chkjaer/main/data/songs_normalize.csv", header = TRUE, stringsAsFactors = FALSE)

all_genres <- as.list(unique(unlist(strsplit(top_spotify_hits2000_2019$genre, ", "))))

create_df <- function(the_genre) {
  new_data <- top_spotify_hits2000_2019 %>%
    filter(str_detect(genre, the_genre)) %>%
    group_by(year) %>%
    summarise(total_popularity_of_genre = sum(popularity, na.rm = TRUE)) %>%
    return(new_data)
}

create_plot_for_genre <- function(new_data) {
  ggplot(new_data, aes(x = year, y = total_popularity_of_genre)) + 
    geom_point() +
    labs(title = "Popularity of Genre from 2000 - 2019",
         x = "Year",
         y = "Total Popularity of Genre")
}

rock_df <- create_df("rock")
create_plot_for_genre(rock_df)

for (genres in all_genres) {
  df <- create_df(genres)
  create_plot_for_genre(df)
}

```

This is an example of how the popularity of rock has decreased over time. You can use the code above to filter select genres such as rock, pop, R&B, etc. to see the popularity over time. Each dot on the chart represents how popular the genre was on the scale used in our data set.

## Chart 3

```{r  echo = FALSE, results = 'asis',  warning=FALSE}
library("tidyverse")
library("ggplot2")

  artists <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/top_artists.csv", stringsAsFactors = FALSE)
  capitals <- read.csv("https://media.githubusercontent.com/media/info201b-au2022/project-Chkjaer/main/data/concap.csv")
top_artists <- artists %>% 
  arrange(-listeners_lastfm) %>%
  head(1000) %>% rename(country = country)

capitals <- rename(capitals, country = CountryName)
  
top_artists <- top_artists %>%
  left_join(capitals, by = "country")

country_shape <- map_data("world")

view(country_shape)

world_map <- ggplot(country_shape) +
  geom_polygon(
    mapping = aes(x = long, y = lat, group = group),
    fill = "grey",
    color = "black",
    size = .01,
  ) +
  coord_map() +
  labs(
    title = "Top 1000 Spotify Artists' Home Countries by Capital and Amount of Listeners",
    x = NULL,
    y = NULL,
  )

world_map + 
  geom_point(
    data = top_artists,
    mapping = aes(x = CapitalLongitude, y = CapitalLatitude, size = listeners_lastfm),
    color = "green",
    alpha = .3
  ) + 
  labs(
    size = "Listeners"
  )
  
```

This chart shows a map of the world where each green dot is placed on the capital of the country. The size of the dot represents how many Spotify listeners each country has, which is scaled through the size of the dot.
